Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex

\begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, for every $\tau\in\Tspace$ and $\vtheta\in\Theta$, there is a positive constant $\Gamma<\infty$ such that:
	\norm[2]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
For REINFORCE we have, from Assumption \ref{ass:bounded_score}:
	\norm[2]{g(\tau\vert\vtheta)}^2 &=
	\norm[2]{\score{\vtheta}{\tau}R(\tau)}^2 \\
	&=\norm[2]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}(a_t | s_t)\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.

>>>>

\begin{restatable}[]{lemma}{bounded-g}\label{lemma:gbound}
Under Assumption \ref{ass:bounded_score}, whether we use the REINFORCE or the G(PO)MDP gradient estimator, for every $\tau\in\Tspace$ and $\vtheta\in\Theta$, there is a positive constant $\Gamma<\infty$ such that:
\[
	\norm[2]{g(\tau\vert\vtheta)}^2 \leq \Gamma.
\]
\end{restatable}
\begin{proof}
For REINFORCE we have, from Assumption \ref{ass:bounded_score}:
\begin{align*}
	\norm[2]{g(\tau\vert\vtheta)}^2 &=
	\norm[2]{\score{\vtheta}{\tau}R(\tau)}^2 \\
	&=\norm[2]{\left(\sum_{t=0}^{H-1}\score{\vtheta}{a_t\vert s_t}\right)\left(\sum_{t=0}^{H-1}\gamma^t r_t\right)}^2\leq H^2G^2\frac{(1-\gamma^H)^2}{(1-\gamma)^2}R^2\dim(\vTheta) \coloneqq \Gamma
\end{align*}
For G(PO)MDP, we do not have a compact expression for $g$, but since it is derived from REINFORCE by neglecting some terms of the kind $\nabla \log \pi_{\vtheta}(a_t | s_t)\gamma^t r_t$ \cite{baxter2001infinite,peters2008reinforcement}, the above bound still holds.
\end{proof}

%New Lemma on Variance
\begin{restatable}[]{lemma}{varineq}\label{lemma:varineq}
For any random vector X, the variance (as defined in Definition \ref{def:var}), can be bounded as follows:
\[
	\Vars{X} \leq \Es{\norm[]{X}^2}.
\]
\end{restatable}
\begin{proof}
By using basic properties of expected value and scalar variance:
\begin{align*}
	\Vars{X} = \Es{\norm[]{X-\Es{X}}^2} &= \Es{\sum_{i=1}^{\dim(X)}\left(X_i-\Es{X_i}\right)^2} = \sum_{i=1}^{\dim(X)}\Es{\left(X_i-\Es{X_i}\right)^2} \\
	&\leq \sum_{i=1}^{\dim(X)}\Es{X_i^2} = \Es{\sum_{i=1}^{\dim(X)}X_i^2} = \Es{\norm[]{X}^2}.
\end{align*}
\end{proof}
